Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2024-09-28 thesis/article.tex
Thesis: Starts Pricing games chapter with 0 sensitivity section, also moves Examples to Appendix

\renewcommand{\maketitle}{
	\begin{titlepage}
		\newgeometry{left=4cm, top=2.3cm, bottom=2.3cm}
		\hbox{\mbox{\hspace{-1.3cm}}
			\vrule depth 0.98\textheight
			\mbox{\hspace{1cm}}
			\vtop{
				\vspace{2cm}
				\begin{flushleft}
					\huge{\bf Profit Games in Heterogeneous 2-link Parallel Networks\\}
					\vskip3cm
					\Large  Thomas Pappas\\
					AL1.18.0011\\
					\vskip3cm
					\begin{minipage}{7.5cm}
						\begin{flushleft}
							\normalsize {\it {\bf Examination committee:}\\
								Dimitris Fotakis, School of Electrical and Computer Engineering, National Technical University of Athens.\\
								Professor's name, Department or School, Institution.\\
								Professor's name, Department or School, Institution.}
						\end{flushleft}
					\end{minipage}
					\hskip0.5cm
					\begin{minipage}{6cm}
						\begin{flushleft}
							\normalsize {\it {\bf Supervisor:}\\
								Dimitris Fotakis, Professor, \\ School of Electrical and Computer Engineering,\\
								National Technical University of Athens.\\
							}
						\end{flushleft}
					\end{minipage}
				\end{flushleft}
				\vskip6cm
				\hskip4.5cm
				\begin{minipage}{5cm}
					\begin{center}
						\includegraphics[width=0.8\textwidth]{alma.png}
					\end{center}
		\end{minipage}}}
	\end{titlepage}
}

@@

@@

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money/time valuation ratios.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of each edge as $c_i^p(t) = \ell_i(x_i(t)) + \alpha(p) \cdot t_i$ and then seek the shortest link among them.
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, \infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
Finally, we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Example \ref{example:a_0_0}, allowing many players close to $0$ breaks the pricing game.

@@

\begin{theorem}
	\label{theorem:alpha_upper_irrelevant}
	Let $([2], \ell, \alpha^{(1)})$ and $([2], \ell, \alpha^{(2)})$ two toll games with same latency functions, where $x_1(0), x_2(0) > 0$ for their common flow $x(0)$ and distribution functions $\alpha_1$ and $\alpha_2$ such that $\alpha^{(1)}(p) = \alpha^{(2)}(p), \forall p: p \le \max\{x_1(0), x_2(0)\}$.
	Then the two games are equivalent.
\end{theorem}

@@

\chapter{Profit maximisation and distribution function}

In this section we will examine how different distribution functions $\alpha$ can affect the profit maximisation game between the toll owners.
[TODO: add framework for examining games]
We will restrict our analysis on distribution functions such that $\alpha(0) > 0$, the reason being that we want to avoid two instances.
The first is to ensure that $\alpha(p) > 0$ when $p > 0$, while the second is that $0$ is not an accumulation point.
[TODO: maybe mention \cite{10.1145/780542.780618} where they also take the first restriction as it creates games with no optimal tolls]
The following two examples demonstrate the issues that might occur if these restrictions are not met.

\begin{example}
	\label{example:a_0_0}
	Consider a $2$-link parallel network pricing game with distribution function $\alpha$ for which it does not hold that $\alpha(p) > 0$ when $p > 0$.
\end{example}

The property not holding for $\alpha$ means that $\exists x_\epsilon > 0 : \alpha(x_\epsilon) = 0 \forall p < x_\epsilon$.
That $x_\epsilon$ flow will completely neglect the tolls and play the congestion game only considering the flow latency, therefore they will always select the fastest edge regardless of the toll price.
This in turn means that any toll owner can exploit it and set their toll to an arbitrary high value $T$ which will force all players with $\alpha(p) > 0$ to leave, thus making their own edge the fastest one.
Even if the edge has a latency function with a high increasing rate, there will be at least a small flow $x_{\epsilon^\prime} \le x_\epsilon$ on that edge.
Therefore the owner can achieve any arbitrary profit $P$ by setting their toll price as $T + P/\epsilon$.
Since both player have that strategy available, we conclude that no equilibrium exists in the pricing game.

\begin{example}
	\label{example:a_0_ap}
	Consider a $2$-link parallel network pricing game with $l_1(x) = x, l_2(x) = 2x$ and distribution function $\alpha(p) = p$.
\end{example}

@@

\paragraph{The bad behaviors} are obviously having divergent or undefined best response functions, which we claim occur only when we allow $0$ to be an accumulation point in $\alpha$, since this allows players with arbitrarily low money-sensitivity to always stay on the expensive edge.
Assume that in Example $\ref{example:a_0_ap}$ we use $\alpha(p) = p + 1$ as a distribution function.
Then we get the following best response functions
\begin{align*}
	B_1(t_2) &=
	\begin{cases}
		\sqrt{5 (3 - t_2)} - (3 - t_2), & t_2 < 6/5 \\
		t_2, & 6/5 \le t_2 \le 3/2 \\
		t_2 + 3 - \sqrt{2 t_2 + 6}, & 3/2 < t_2 < 5 \\
		t_2 - 1 & t_2 \ge 5
	\end{cases}\\
	B_2(t_1) &=
	\begin{cases}
		2 \sqrt{3 - t_1} - (3 - t_1), & t_1 \le \frac29 \sqrt{22 - 4 \sqrt{10}} \\
		t_1 + 3 - \frac{\sqrt{10}}2 \sqrt{t_1 + 3}, & \frac29 \sqrt{22 - 4 \sqrt{10}} < t_1 < 7 \\
		t_1 - 2, & t_1 \ge 7
	\end{cases}
\end{align*}
which can be shown [TODO: currently done experimentally in python] that has a $\NE$ at $t \approx (1.1464, 0.9268)$.

@@

\subparagraph{Combining} $B_1^>(t_2), B_1^<(t_2)$ and $B_2^>(t_1), B_2^<(t_1)$ the same way as in Example $\ref{example:a_0_ap}$ we get the final best response functions.
\begin{equation*}
	\begin{aligned}
		B_1(t_2) =
		\begin{cases}
			\frac{t_2 + 2}2, & t_2 \le 1 \\
			t_2 + \frac12, & 1 < t_2 < \frac32 \\
			t_2, & \frac32 \le t_2 \le 2 \\
			\frac{t_2 + 2}2, & 2 < t_2 < 4 \\
			t_2 - 1, & t_2 \ge 4
		\end{cases}
	\end{aligned}
	\qquad
	\begin{aligned}
		B_2(t_2) =
		\begin{cases}
			\frac{t_1 + 1}2, & t_1 \le \frac15 \\
			\frac{5 t_1 + 1}{10}, & \frac15 < t_1 < frac25 \\
			t_1 - \frac1{10}, & \frac25 \le t_1 \le \frac{10 + 2 \sqrt{15}}5 \\
			\frac{t_1 + 1}2, & \frac{10 + 2 \sqrt{15}}5 < t_1 < 5 \\
			t_1 - 2, & t_1 \ge 5
		\end{cases}
	\end{aligned}
\end{equation*}

@@

\section{Equilibrium}

In this section we will show that if the distribution function has the following properties
\begin{itemize}
	\item $\alpha(0) > 0$
	\item $\alpha$ is onto
\end{itemize}
Then a $\NE$ for the pricing game exists ???

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Bibliography}

\bibliographystyle{plain}
\bibliography{ref.bib}

>>>>

\renewcommand{\maketitle}{
	\begin{titlepage}
		\newgeometry{left=4cm, top=2.3cm, bottom=2.3cm}
		\hbox{\mbox{\hspace{-1.3cm}}
			\vrule depth 0.98\textheight
			\mbox{\hspace{1cm}}
			\vtop{
				\vspace{2cm}
				\begin{flushleft}
					\huge{\bf Pricing Games in Heterogeneous 2-link Parallel Networks\\}
					\vskip3cm
					\Large  Thomas Pappas\\
					AL1.18.0011\\
					\vskip3cm
					\begin{minipage}{7.5cm}
						\begin{flushleft}
							\normalsize {\it {\bf Examination committee:}\\
								Dimitris Fotakis, School of Electrical and Computer Engineering, National Technical University of Athens.\\
								Professor's name, Department or School, Institution.\\
								Professor's name, Department or School, Institution.}
						\end{flushleft}
					\end{minipage}
					\hskip0.5cm
					\begin{minipage}{6cm}
						\begin{flushleft}
							\normalsize {\it {\bf Supervisor:}\\
								Dimitris Fotakis, Professor, \\ School of Electrical and Computer Engineering,\\
								National Technical University of Athens.\\
							}
						\end{flushleft}
					\end{minipage}
				\end{flushleft}
				\vskip6cm
				\hskip4.5cm
				\begin{minipage}{5cm}
					\begin{center}
						\includegraphics[width=0.8\textwidth]{alma.png}
					\end{center}
		\end{minipage}}}
	\end{titlepage}
}

@@

\cleardoublepage

@@

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money/time valuation ratios.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of each edge as $c_i^p(t) = \ell_i(x_i(t)) + \alpha(p) \cdot t_i$ and then seek the shortest link among them.
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, \infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
Finally, we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Lemma \ref{lemma:a_0_0}, allowing many players close to $0$ breaks the pricing game.

@@

\begin{theorem}
	\label{theorem:alpha_upper_irrelevant}
	Let $([2], \ell, \alpha^{(1)})$ and $([2], \ell, \alpha^{(2)})$ two toll games with same latency functions, where $x_1(0), x_2(0) > 0$ for their common flow $x(0)$ and distribution functions $\alpha_1$ and $\alpha_2$ such that $\alpha^{(1)}(p) = \alpha^{(2)}(p), \forall p: p < \max\{x_1(0), x_2(0)\}$.
	Then the two games are equivalent.
\end{theorem}

@@

\cleardoublepage

\chapter{Equilibria in Pricing Games}

In this section we introduce the pricing game to our model and examine how different distribution functions $\alpha$ can affect the profit maximisation game between the toll owners, specifically when $\alpha$ is fixed, step or continuous.
We begin, though, with some necessary restrictions.

\section{Problems with $0$ money sensitivity}

In general having players with $\alpha(p) = 0$, i.e. players completely indifferent to tolls, creates issues.
Even Cole et al. \cite{10.1145/780542.780618} assume $\alpha(p) > 0$ when $p > 0$ for their model to support optimal tolls.
In fact, this restriction proves necessary for our pricing game as well, since its absence allows distribution functions with arbitrarily small flow all at $0$, which in turn, as we'll see in the following Lemma, breaks the pricing game, too.

\begin{lemma}
	\label{lemma:a_0_0}
	Let $([2], \ell, \alpha)$ a pricing game for which it holds that $\exists \epsilon > 0: \alpha(p) = 0 \; \forall p < \epsilon$.
	Then the best response functions $B_i(t_{-i})$ are unbounded.
\end{lemma}

\begin{proof}
	All players $p$ with $\alpha(p) = 0$ will completely neglect the toll values and play the congestion game only considering the flow latencies, thus always selecting the fastest edge, regardless of the toll prices.

	We have from assumption that $x_i(0) > 0$.
	Consider $x_i^\prime = \min\{x_i(0), \epsilon\} > 0$.
	Obviously $\alpha(p) = 0 \; \forall p < x_i$.
	Then we can see that for all $t_i > t_{-i}$ where edge $i$ has the lower latency edge, all those players will be on that edge and will remain so regardless of how high is the value of $t_i$.
	For the respective profit that means $\Pi_i(t_i, t_{-i}) \ge x_i t_i \; \forall t_i > t_{-i}$.
	Consequently, as $t_i \rightarrow +\infty$ then so does $x_i t_i$ and $\Pi_i(t_i, t_{-i})$, making in turn $B_i(t_{-i}) \rightarrow +\infty$.
\end{proof}
We therefore need the restriction of $\alpha(p) > 0$ when $p > 0$ in order for our pricing game to have an equilibrium.

\section{Fixed distribution function}

\section{Step distribution function}

\section{Continuous distribution function}

For the continuous distribution function case, the previous restriction of $\alpha(p) > 0$ when $p > 0$ is not sufficient, as it still allows for distribution functions with many player values close to $0$ < Is that actually the problem?.
Specifically the case of $\alpha(p) = cp$, for some $c > 0$, shows some interesting behavior, as it creates instances where $\Pi_i$ becomes either unbounded or fixed; in the first case then $B_i$ becomes unbounded, while in the second it becomes an an infinite set.
This is showcased analytically in Example \ref{example:ap_p} in the Appendix, where for $\alpha(p) = p$ we see that the resulting best responses can indeed be undefined or unbounded.

When does this happen?
Is the problem $0$ being an accumulation point?
\begin{lemma}[Conjecture]
	If a distribution function has $\alpha(p) \le p \; \forall p$, then $\Pi_i$ becomes either unbounded or fixed.
\end{lemma}

We will avoid all the above bad behaviors by assuming in all further cases that $\alpha(0) > 0$.

\cleardoublepage

\chapter{Future work}

\cleardoublepage

\phantomsection
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{plain}
\bibliography{ref.bib}
\cleardoublepage

\phantomsection
\addcontentsline{toc}{chapter}{Appendix}
\chapter*{Appendix}

[TODO: add framework for examining games]

\begin{example}
	\label{example:ap_p}
	Consider a $2$-link parallel network pricing game with $l_1(x) = x, l_2(x) = 2x$ and distribution function $\alpha(p) = p$.
\end{example}

@@

\paragraph{The bad behaviors} are obviously having divergent or undefined best response functions, which we claim occur only when we allow $0$ to be an accumulation point in $\alpha$, since this allows players with arbitrarily low money-sensitivity to always stay on the expensive edge.
Assume that in Example $\ref{example:ap_p}$ we use $\alpha(p) = p + 1$ as a distribution function.
Then we get the following best response functions
\begin{align*}
	B_1(t_2) &=
	\begin{cases}
		\sqrt{5 (3 - t_2)} - (3 - t_2), & t_2 < 6/5 \\
		t_2, & 6/5 \le t_2 \le 3/2 \\
		t_2 + 3 - \sqrt{2 t_2 + 6}, & 3/2 < t_2 < 5 \\
		t_2 - 1 & t_2 \ge 5
	\end{cases}\\
	B_2(t_1) &=
	\begin{cases}
		2 \sqrt{3 - t_1} - (3 - t_1), & t_1 \le \frac29 \sqrt{22 - 4 \sqrt{10}} \\
		t_1 + 3 - \frac{\sqrt{10}}2 \sqrt{t_1 + 3}, & \frac29 \sqrt{22 - 4 \sqrt{10}} < t_1 < 7 \\
		t_1 - 2, & t_1 \ge 7
	\end{cases}
\end{align*}
which can be shown [TODO: currently done experimentally in python] that has a $\NE$ at $t \approx (1.1464, 0.9268)$.

@@

\subparagraph{Combining} $B_1^>(t_2), B_1^<(t_2)$ and $B_2^>(t_1), B_2^<(t_1)$ the same way as in Example $\ref{example:ap_p}$ we get the final best response functions.
\begin{equation*}
	\begin{aligned}
		B_1(t_2) =
		\begin{cases}
			\frac{t_2 + 2}2, & t_2 \le 1 \\
			t_2 + \frac12, & 1 < t_2 < \frac32 \\
			t_2, & \frac32 \le t_2 \le 2 \\
			\frac{t_2 + 2}2, & 2 < t_2 < 4 \\
			t_2 - 1, & t_2 \ge 4
		\end{cases}
	\end{aligned}
	\qquad
	\begin{aligned}
		B_2(t_2) =
		\begin{cases}
			\frac{t_1 + 1}2, & t_1 \le \frac15 \\
			\frac{5 t_1 + 1}{10}, & \frac15 < t_1 < frac25 \\
			t_1 - \frac1{10}, & \frac25 \le t_1 \le \frac{10 + 2 \sqrt{15}}5 \\
			\frac{t_1 + 1}2, & \frac{10 + 2 \sqrt{15}}5 < t_1 < 5 \\
			t_1 - 2, & t_1 \ge 5
		\end{cases}
	\end{aligned}
\end{equation*}

@@
